Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-02-14 3dpas_intro.tex
light edits in intro and conclusions

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  Occasionally data are often
naturally distributed and have to be processed near source; sometimes data that
are localized have to be distributed in order to be processed, either for
reasons of performance or for collaborative analytics. Further, the scale of
varying data source and destination, as well as the transfer of production rates
makes temporal variations in data properties increasingly important.  With an
increase in the importance of distributed and temporal properties of data,
systems and infrastructure aspects such as resource scheduling, data placement,
and transfer decisions that were statically determinable at small scales, emerge
as dynamic and important considerations at large scales of data and
distribution.

@@

The analytical space spanned by the three ``D''s of data-intensive, distributed,
and dynamic provides a common vocabulary and effective framework to analyze
applications and infrastructure alike.  Subjecting a range of otherwise distinct
applications to a common analytical framework provides an opportunity to search
for similarities across applications and understand core differences.  The
diverse D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.  A common framework thus informs infrastructure scientists and
cyberinfrastructure experts where effort is likely to yield maximum returns.
The advantages of providing a common but extensible vocabulary that can be used
by the entire community, application end-users, developers and infrastructure
scientists alike, in an emerging field cannot be overstated. We believe that
such vocabulary and terminology has been missing from the practice of
distributed cyberinfrastructure in general, with negative consequences.

@@

One motivation of this work is to understand the dynamic properties
and the distribution of applications, and the complex interplay
between them that arises as a consequence of scale.  Additionally, we
are motivated by an attempt to understand infrastructure---existing
capabilities, trends, and limitations---available to the complexity,
challenges and characteristics of dynamic and distributed data at
scale.

We would like the reader to keep in mind some context and scope-setting points:

@@

\item This {\it survey} is not meant to be complete, either in scope of domains
  covered or types of infrastructure and application surveyed.  They
  represent the interests and experience of the authors. The specific
  requirements and barriers of applications cited in this paper are
  representative of the time at which the applications were surveyed
  (2012-2013); this is also true of the quantitative analysis of applications.

@@

\item The focus is on applications and infrastructure that arise from science
  and engineering projects in academia; this is not exclusive but is more a
  guiding principle.  While we are cognizant of enterprise applications and
  infrastructure, and recognize the impressive advances made therein, we also
  acknowledge that the factors such as scale, distribution, open-source, and
  interoperability influence the solutions employed and infrastructures used; it
  is important to therefore focus on problems that emerge as in the space
  defined by these ``academic constraints''.

@@

\item Distribution can occur on many levels.  In this paper, we use {\em locally
    distributed} to indicate that data/compute resides on multiple nodes within
  one data center.  {\em Geographically distributed} describes the distribution
  of data and/or compute across multiple data centers.  Also, our focus is
  primarily on infrastructure considerations at the large-scale, as searching
  for common solutions and patterns at the large-scale is likely to be more
  fruitful than at small-scales where customized tools and optimization of
  low-level systems features is likely to return greater yield.

@@

The rest of this paper consists of
the application scenarios (\S\ref{sec:scenarios}),
a discussion of the aspects of distribution and dynamism in the
applications (\S\ref{sec:distdyndata}),
a description of the software infrastructure required for D3 applications
% Yogesh: This section's description seemed more verbose than others, hence commenting out details
% , such as
% software frameworks, tools, and middleware and platform services
(\S\ref{sec:infrastructure}),
analysis and characterization of the applications and
conclusions (\S\ref{sec:conclusions}).

@@

The growth in data volumes and its importance is having profound implications on
the way applications are designed to the way we develop infrastructure and
provision associated services.  Many fields, such as biology, that until
recently were not characterized by their intensive use of data are rapidly
becoming data-driven.

@@

An important motivation for this work is to capture the state of the art in the
design and development of distributed dynamic applications and the associated
infrastructure currently used for their execution. We have aimed to try to
discern common approaches and challenges, and to identify any obvious gaps (or
opportunities). Using a common vocabulary and terminology to describe otherwise
distinct and unrelated applications allows potential application
users/developers to benefit from the insight that a common analytical framework
provides. For example, seeing how applications with similar patterns (e.g.,
pipelines) have addressed issues of dynamism and distribution enables common
solutions and facilitates the emergence of best practices even in the absence of
formal methods.  Similarly, our analysis sends a useful message to domain and
application scientists that the challenges and barriers they face are not unique
even if current solutions require a level of customization.  In this paper, we
have identified many different applications characteristics, programming
systems, and infrastructure techniques that either support dynamic data or
produce it.

>>>>

The landscape of scientific and enterprise computing is being fundamentally
altered by prodigious volumes of data.  Occasionally data are often
naturally distributed and have to be processed near their source, and sometimes data that
are localized have to be distributed in order to be processed, either for
reasons of performance or for collaborative analytics. Further, the scale of
varying data source and destination, as well as the transfer of production rates \katznote{what are `transfer of production rates'?}
makes temporal variations in data properties increasingly important.  With an
increase in the importance of distributed and temporal properties of data,
systems and infrastructure aspects such as resource scheduling, data placement,
and transfer decisions that were statically determinable at small scales emerge
as dynamic and important considerations at large scales of data and
distribution.

@@

The analytical space spanned by the three ``D''s of dynamic, distributed,
and data-intensive provides a common vocabulary and effective framework to analyze
applications and infrastructure alike.  Subjecting a range of otherwise distinct
applications to a common analytical framework provides an opportunity to search
for similarities across applications and understand core differences.  The
diverse D3 scenarios point to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.  A common framework thus informs infrastructure scientists and
cyberinfrastructure experts where effort is likely to yield maximum returns.
The advantages of providing a common but extensible vocabulary that can be used
by the entire community---application end-users, developers and infrastructure
scientists alike---in an emerging field cannot be overstated. We believe that
such a vocabulary and terminology has been missing from the practice of
distributed cyberinfrastructure in general, with negative consequences.

@@

One motivation of this work is to understand the dynamic properties
and the distribution of applications, and the complex interplay
between them that arises as a consequence of scale.  Additionally, we
are motivated by an attempt to understand infrastructure---existing
capabilities, trends, and limitations---available to support the complexity,
challenges and characteristics of dynamic and distributed data at
scale.

We would like the reader to keep in mind some points regarding context and scope:

@@

\item This {\it survey} is not meant to be complete, either in the scope of domains
  covered or the types of infrastructure and application surveyed.  It
  represent the interests and experience of the authors. The specific
  requirements and barriers of applications cited in this paper are
  representative of the time at which the applications were surveyed
  (2012-2013); this is also true of the quantitative analysis of applications.

@@

\item The focus is generally on applications and infrastructure that arise from science
  and engineering projects in academia in general; this is a
  guiding principle.  While we are cognizant of enterprise applications and
  infrastructure, and recognize the impressive advances made therein, we also
  acknowledge that the factors such as scale, distribution, open-source, and
  interoperability influence the solutions employed and infrastructures used; it
  is important to therefore focus on problems that emerge in the space
  defined by these ``academic constraints''.

@@

\item Distribution can occur on many levels.  In this paper, we use {\em locally
    distributed} to indicate that data/compute resides on multiple nodes within
  one data center.  {\em Geographically distributed} describes the distribution
  of data and/or compute across multiple data centers.  Also, our focus is
  primarily on infrastructure considerations at  large scale, as searching
  for common solutions and patterns at  large scale is likely to be more
  fruitful than at small scales where customized tools and optimization of
  low-level systems features is likely to return greater yield.

@@

The remainder of this paper consists of
the application scenarios (\S\ref{sec:scenarios}),
a discussion of the aspects of distribution and dynamism in the
applications (\S\ref{sec:distdyndata}),
a description of the software infrastructure required for D3 applications
% Yogesh: This section's description seemed more verbose than others, hence commenting out details
% , such as
% software frameworks, tools, and middleware and platform services
(\S\ref{sec:infrastructure}),
analysis and characterization of the applications and
conclusions (\S\ref{sec:conclusions}).

@@

The growth in data volumes and its importance is having profound implications on
the way applications are designed and the way we develop infrastructure and
provision associated services.  Many fields, such as biology, that until
recently were not characterized by their intensive use of data are rapidly
becoming data-driven.

@@

An important motivation for this work is to capture the state of the art in the
design and development of distributed dynamic applications and the associated
infrastructure currently used for their execution. We have aimed to try to
discern common approaches and challenges, and to identify any obvious gaps (or
opportunities). Using a common vocabulary and terminology to describe otherwise
distinct and unrelated applications allows potential application
users\slash developers to benefit from the insights that a common analytical framework
provides. For example, seeing how applications with similar patterns (e.g.,
pipelines) have addressed issues of dynamism and distribution enables common
solutions and facilitates the emergence of best practices even in the absence of
formal methods.  Similarly, our analysis sends a useful message to domain and
application scientists that the challenges and barriers they face are not unique
even if current solutions require a level of customization.  In this paper, we
have identified many different applications characteristics, programming
systems, and infrastructure techniques that either support dynamic data or
produce it.
